---
title: "Data606 project proposal"
author: "angel"
date: "2024-04-04"
output:
  pdf_document: default
  html_document: default
---


### Data Preparation


```{r}
library(ggplot2)
library(dplyr)
library(scales)
library(tidyverse)
```

```{r setup, echo=TRUE, results='hide', warning=FALSE, message=FALSE}
# load data
df <- read.csv("https://raw.githubusercontent.com/Angelogallardo05/Data606-proposal/main/NY-House-Dataset.csv")
```

```{r}
df <- na.omit(df)


```

### Research question 

**Can the amount of square footage, rooms, bathrooms, longitude, and latitude predict a home price in NYC? **



### Cases 

**There is data on about 5K homes for sale in NYC **



### Data collection 

**Kaggle**



### Type of study 

**the effect of sq footage, beds, bathrooms, in home prices.**



### Data Source 

**https://www.kaggle.com/datasets/nelgiriyewithana/new-york-housing-market**




### Dependent Variable

**home market price, quantitative**



### Independent Variable(s)

**square footage, bathrooms, rooms. borough**


### Relevant summary statistics 

**Provide summary statistics for each the variables. Also include appropriate visualizations related to your research question (e.g. scatter plot, boxplots, etc). This step requires the use of R, hence a code chunk is provided below. Insert more code chunks as needed.**


```{r}
library(dplyr)
df_cleaned <- df %>%
  na.omit() %>%
  filter(PROPERTYSQFT != 2184.207862) %>%
   filter(!is.na(PRICE), !grepl("[^0-9.]", PRICE)) %>%
  mutate(PRICE = as.numeric(PRICE))
```



```{r}

df_grouped <- df_cleaned %>%
  filter(TYPE %in% c("House for sale", "Condo for sale", "Townhouse for sale", "Multi-family home for sale")) %>%
  select(TYPE, PRICE, BEDS, BATH, PROPERTYSQFT, LONGITUDE, LATITUDE) %>%
  na.omit()


ggplot(df_grouped, aes(x = PROPERTYSQFT, y = PRICE)) +
  geom_point() +  
  geom_smooth(method = "lm", se = FALSE) +  
  labs(x = "Property Square Footage", y = "Price", title = "Price vs. Property Square Footage by Type") +
  scale_y_continuous(labels = scales::comma_format(scale = 1e-3)) +
  facet_wrap(~ TYPE, scales = "free") +  
  theme_minimal()


ggplot(df_grouped, aes(x = BEDS, y = PRICE)) +
  geom_point() +  
  geom_smooth(method = "lm", se = FALSE) +  
  labs(x = "Beds", y = "Price", title = "Price vs. Beds by Type") +
  scale_y_continuous(labels = scales::comma_format(scale = 1e-3)) +
  facet_wrap(~ TYPE, scales = "free") +
  theme_minimal()


ggplot(df_grouped, aes(x = BATH, y = PRICE)) +
  geom_point() +  
  geom_smooth(method = "lm", se = FALSE) +  
  labs(x = "Baths", y = "Price", title = "Price vs. Baths by Type") +
  scale_y_continuous(labels = scales::comma_format(scale = 1e-3)) +
  facet_wrap(~ TYPE, scales = "free") +  
  theme_minimal()

ggplot(df_grouped, aes(x = LATITUDE, y = PRICE)) +
  geom_point() +  
  geom_smooth(method = "lm", se = FALSE) +  
  labs(x = "Latitude", y = "Price", title = "Price vs. Latitude by Type") +
  scale_y_continuous(labels = scales::comma_format(scale = 1e-3)) +
  facet_wrap(~ TYPE, scales = "free") +  
  theme_minimal()


ggplot(df_grouped, aes(x = LONGITUDE, y = PRICE)) +
  geom_point() +  
  geom_smooth(method = "lm", se = FALSE) +  
  labs(x = "Longitude", y = "Price", title = "Price vs. Longitude by Type") +
  scale_y_continuous(labels = scales::comma_format(scale = 1e-3)) +
  facet_wrap(~ TYPE, scales = "free") +  
  theme_minimal()

```




```{r}
selected_cols <- df_cleaned %>%
  filter(TYPE == 'House for sale') %>%
  select(BEDS, BATH, PROPERTYSQFT, LATITUDE, LONGITUDE) %>%
  na.omit()  


selected_cols_long <- selected_cols %>%
  pivot_longer(cols = c(BEDS, BATH, PROPERTYSQFT, LATITUDE, LONGITUDE),
               names_to = "Variable", values_to = "Value")


ggplot(selected_cols_long, aes(x = Value, fill = Variable)) +
  geom_histogram(bins = 50, alpha = 0.7, position = "identity") +
  facet_wrap(~ Variable, scales = "free") +
  labs(title = "Histogram of Numeric Variables",
       x = "Value", y = "Frequency") +
  theme_minimal()
  
```






